Automated Formal Synthesis of Provably Safe Digital Controllers for Continuous Plants

### 2019-01-31 automatica/paper.tex
some responses

In addition to precisely replicating the finite-precision arithmetic of the digital controller, 
we must consider that our \emph{model} itself in \eqref{eq:plant} employs finite-precision arithmetic to represent the behaviour of the real system. 
In order to guarantee soundness, we therefore encompass the error that is due to modelling (as opposed to the nature of the digital controller):   
the representations used in the plant model and its arithmetic operations are carried out at finite precision.  
\begin{itemize} 
\item
We account for the error introduced by finite precision arithmetic applied over the plant variables $x_k$ and $u_k$ \note{here I have replaced the continuous-time signals with discrete-time ones}, which are actually continuous quantities. 
We guarantee that the precision we use to represent the plant variables is at least as precise as the precision used in the digital controller, 
and we use interval arithmetic to bound the incurred errors, as further detailed in Section~\ref{sec:numeric_rep}. 
\end{itemize}

@@

We fix an index $k$, 
and we synthesize a safe controller by
unfolding the transition system (.e., the closed-loop model) $k$ steps and by selecting a controller $K$ and a single initial state, 
such that the states at each step do not violate the safety criteria (see Section~\ref{sec:safety}).    
That is, we ask the bounded model checker~\cite{ClarkeKL04} if there exists a $K$ that is safe for at least one $x_0$ in our set of all possible initial states, and a given fixed-point precision for the controller, and where the input signal remains within the specified bounds. 
This approach is sound \note{you mean complete?} if the current $k$ is greater than the completeness threshold (see later step).  
We also assume some finite precision for the plant and a given sampling time \note{sampling time comes out of nowhere - also not in the algorithm}.  
We use a fixed-point precision in this description, given by $\langle I_p,F_p\rangle$, but if we are considering a floating-point controller we will instead model the plant with a floating-point precision, $\langle E_p,M_p\rangle$.
The checks that these assumptions hold are performed by the subsequent {\sc verify} stages.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.9\textwidth]{figures/synthesis-flow.pdf}
\caption{Phases of the controller synthesis}
\label{fig:synthesis-flow}
\end{figure*}

The implementation of the proposed methodology for synthesis of digital controls for physical plants is based on the Digital-System Synthesizer 
(DSSynth) tool~\cite{DBLP:conf/kbse/AbateBCCCDKKP17}, 
which can be split into two main stages: manual and automated, as illustrated in Fig.~\ref{fig:synthesis-flow} and detailed next.

In Step~1, the user selects the system representation, which can be a transfer function or a state-space model.  
In Step~2, the plant (e.g., in the form of Equation~\eqref{eq:ode1}) is provided~\cite{astrom1997computer}.  
Finally, in Step~3, the numerical representation for the digital controller implementation must be set by the user, i.e., 
the finite precision arithmetic that defines the number of bits of integer and fractional parts when using fixed-point arithmetic, 
or half- and single-precision when using floating-point arithmetic, 
as well as the input range.  
% This input, provided by the user, is a specification of the physical plant.

In the second stage, 
the automated synthesis process starts with Step~A, where the
\tool~translates the model specification into an ANSI-C program.  
In Step~B, the proposed CEGIS engine is invoked to synthesize the digital controller w.r.t.~the specification on the digital model.  
Finally, in Step~C, the synthesized digital controller is generated.  
The output by \tool~is the synthesized digital controller represented either as a transfer function or in state-space form.  
The synthesis is considered to be \emph{successful} if a digital controller is
correctly synthesized w.r.t. effects of finite precision arithmetic; or if there is no solution (no
controller) for the corresponding physical plant (given the implementation
requirements), then the synthesis is considered to have \emph{failed} \note{we don't seem to have any such instance.}.

Our CEGIS engine is implemented as an integrated module within the C bounded
model checker (CBMC)~\cite{ClarkeKL04}.  CBMC transforms the ANSI-C representation
of the closed-loop control model model into its internal representation (IR).  We
instrument this IR for each synthesis or verification scenario accordingly,
and use CBMC as an oracle to answer our queries.  CBMC itself relies on an
underlying SAT or SMT solver to address verification conditions.  We~model
effects of finite precision arithmetic 
explicitly using CBMC's nondeterminism API (e.g., \texttt{nondet},
\texttt{CPROVER\_assume} intrinsic functions).  Our module is included into
CBMC~5.7 and is available for
download.\footnote{\url{https://github.com/diffblue/cbmc/archive/cbmc-5.7.zip}}
We further created a VirtualBox OVA image with our tool pre-installed,
including our MATLAB-generated benchmarks and benchmark shell script with
instructions to run all benchmarks and to reproduce our
experiments.\footnote{\note{\url{XXXXXXXXXX}}}
%
%\textcolor{red}{This paragraph needs to be updated by Elizabeth and Dario.}
%
%The entire implementation consists of $XXX$ and $XXX$ lines of C code in the multi-stage and abstract
%accelerator back-ends, respectively, $XXXX$ lines of C code for utility functions such as interval, 
%fixed- and floating-point arithmetic, as well as $1000$ lines of C++ code in the control synthesis
%module in CEGIS CBMC.  We evaluate \tool~using $XXX$ SISO control system
%benchmarks.

@@

Our benchmark suite consists of 19 SISO case studies extracted from the 
control literature~\cite{acrobot,cstr,CHEN1979389,KOKOTOVIC198023,gajic2008optimal,Franklin15, maglev, converters, CTMS, gajic2008optimal}. 
The models are discretized with sampling times ranging from 1s to 0.0001s. CEGIS initially starts with the coarsest discretisation (i.e., 1s sampling time), and if it fails to find a controller, reduces the sampling time.

@@

that is a computer storage device used to read and write data on magnetic tapes.

The \textit{USCG cutter tampa heading angle} plant describes the heading angle dynamics of a Coast 
Guard cutter.

@@

Using the state-space models described in Section~\ref{exp:benchmarks}, 
the evaluation study has the following experimental goal:

\item[EG1] \textbf{(Performance)} Show that the multi-staged 
%and the abstraction-based 
CEGIS approach is able to generate finite precision digital controllers using fixed-point and floating-point arithmetic in a reasonable amount of time.

%\item[EG2] \textbf{(Sanity check)} Confirm the stability \note{also?} and safety of the synthesized controllers outside 
%of our model representation using MATLAB. 
%\note{Where is this done?}

@@

We provide the results in Table~\ref{tab:results}, where: 
\textit{Benchmark} is the name of the corresponding benchmark; 
\textit{Order} is the number of continuous variables of the model; 
$\mathcal{F}_{\langle I_p,F_p \rangle}$, where $I_p$ and $F_p$ indicate the integer and fractional parts,  
and $\mathcal{F}_{\langle E_p,M_p \rangle}$, where $E_p$ and $M_p$ indicate the exponent and mantissa part, 
are the fixed- and floating-point precisions used to model the plant, respectively; 
and \textit{Time} is the total time required to synthesize a controller for the given plant.    
Timeouts are indicated by~\xmark.  
The precision for the controller, $\mathcal{F}_{\langle I_c,F_c \rangle}$, is
chosen to be $I_c = 8$, $F_c = 8$ for fixed-point, whereas  
$\mathcal{F}_{\langle E_c, M_c \rangle}$ is chosen to be $W_c = 6$ and $F_c = 10$
for half-precision floating-point format.

@@

The mean run-time for the successful benchmarks is 225\,s for fixed-point and 125\,s for floating-point implementations, respectively.  
The median run-time for the successful benchmarks is 15.7\,s for fixed-point and 17.8\,s for floating-point implementations, respectively.
We consider these times to be short enough to be of practical use to control engineers, and thus affirm the validity of EG1.

We present in this paper a selection of case studies for which we are able to synthesise controllers. In the full experimental environment, 
which is available to download, we include further case studies that we were unable to synthesise controllers for. This is in many cases
due to timeouts, especially when the completeness threshold is too large, 
or a controller simply may not exist for a benchmark at a given time discretisation and required controller precision. 
Yet another source of incompleteness is the inability of the {\sc synthesize} phase to use a large enough precision for the plant model.

@@

@@

\textit{Benchmark selection:} We report an assessment of both our approaches
over a diverse set of real-world benchmarks.  Nevertheless, this set of
benchmarks is limited within the scope of this paper and the performance may
not generalize to other benchmarks.
%\note{is this where we attempt asserting EG2?}

\textit{Plant precision and discretization heuristics:} Our algorithm to
select suitable finite precision arithmetic to model the plant behavior 
increases the precision by 8 bits at each step 
in order to be compliant with the CBMC type API. 
Similarly, for time-discretization, we try a set of pre-defined time discretisations.
This works sufficiently well for our benchmarks, but
performance may suffer in some cases, for example if the completeness
threshold is high.

@@

%\note{So, what about EG2? Shall we assert it? }

@@

We have presented an automated approach to synthesize digital
state-feedback controllers that ensure both stability and safety over state-space models.   
%The first approach relies on unfolding of the closed-loop model dynamics up to a completeness threshold, while the second
%one applies abstraction-refinement and acceleration to increase speed, whilst retaining soundness.
%
Our approach is novel within the digital control literature: we provide a fully
automated synthesis method that is algorithmically and numerically sound,
considering various error sources in the implementation of the digital
control algorithm and in the computational modeling of plant dynamics.
%
Our experimental results show we are able to synthesize fully automatically 
safe controllers for diverse benchmarks from the control literature, 
within a reasonable amount of time.

%
%\note{Let us collect here the main extensions we regard to be relevant: \\ 
%models with partial observations; \\ 
%implementation with SMT; \\ 
%continuous-time; \\ 
%more complex specs; \\ 
%\ldots }

>>>>

In addition to precisely replicating the finite-precision arithmetic of the digital controller, 
we must consider that our \emph{model} itself in \eqref{eq:plant} employs finite-precision arithmetic to represent the behaviour of the real system. 
In order to guarantee soundness, we therefore encompass the error that is due to modelling (as opposed to the nature of the digital controller):   
the representations used in the plant model and its arithmetic operations are carried out at finite precision.  
\begin{itemize} 
\item
We account for the error introduced by finite precision arithmetic applied over the plant variables $x_k$ and $u_k$, 
%\note{here I have replaced the continuous-time signals with discrete-time ones}, 
which are actually continuous quantities. 
We guarantee that the precision we use to represent the plant variables is at least as precise as the precision used in the digital controller, 
and we use interval arithmetic to bound the incurred errors, as further detailed in Section~\ref{sec:numeric_rep}. 
\end{itemize}

@@

We fix an index $k$, 
and we synthesize a safe controller by
unfolding the transition system (.e., the closed-loop model) $k$ steps and by selecting a controller $K$ and a single initial state, 
such that the states at each step do not violate the safety criteria (see Section~\ref{sec:safety}).    
That is, we ask the bounded model checker~\cite{ClarkeKL04} if there exists a $K$ that is safe for at least one $x_0$ in our set of all possible initial states, and a given fixed-point precision for the controller, and where the input signal remains within the specified bounds. 
This approach is sound, i.e., the controller produced is guaranteed to be safe, \note{you mean complete? No, i mean sound} if the current $k$ is greater than the completeness threshold (see later step).  
We also assume some finite precision for the plant and a given time discretisation, as described in section~\ref{sec:model}.  
We use a fixed-point precision in this description, given by $\langle I_p,F_p\rangle$, but if we are considering a floating-point controller we will instead model the plant with a floating-point precision, $\langle E_p,M_p\rangle$.
The checks that these assumptions hold are performed by the subsequent {\sc verify} stages.

@@

\begin{figure*}[t]
\centering
\includegraphics[width=0.9\textwidth]{figures/synthesis-flow.pdf}
\caption{The Digital-System Synthesizer (\tool) tool - distinct phases of the controller synthesis }
\label{fig:synthesis-flow}
\end{figure*}

The implementation of the proposed methodology for synthesis of digital controls for physical plants is based on the Digital-System Synthesizer 
(\tool) tool~\cite{DBLP:conf/kbse/AbateBCCCDKKP17}, 
which can be split into two main stages: manual and automated, as illustrated in Fig.~\ref{fig:synthesis-flow} and detailed next.

The first stage comprises the following steps. 
In Step~1, the user selects the system representation, which can be a transfer function or a state-space model (we focus on the latter in this article).   
In Step~2, the plant (e.g., in the form of Equation~\eqref{eq:ode1}) is provided~\cite{astrom1997computer}.  
Finally, in Step~3, the numerical representation for the digital controller implementation must be set by the user: 
this is the finite-precision arithmetic that defines the number of bits of integer and fractional parts when using fixed-point arithmetic, 
or half- and single-precision when using floating-point arithmetic. 
The user also specifies the input range.

In the second stage, 
the automated synthesis process starts with Step~A, where the
\tool~translates the model specification into an ANSI-C program.  
In Step~B, the discussed CEGIS engine is invoked, 
in order to synthesize the digital controller w.r.t.~the specification given on the model.  
Finally, in Step~C, the synthesized digital controller is generated.  
The output of \tool~is the synthesized digital controller \note{you mean: the closed-loop model}, 
which is represented either as a transfer function or in state-space form (in this article we consider the latter case).  
The synthesis is considered to be \emph{successful} if a digital controller is
correctly synthesized with respect to the effects of finite-precision arithmetic; 
otherwise, if there is no solution (no controller) for the corresponding physical plant (given the implementation requirements), 
then the synthesis is considered to have \emph{failed}.

Our CEGIS engine is implemented as an integrated module within the C bounded
model checker (CBMC)~\cite{ClarkeKL04}.  CBMC transforms the ANSI-C representation
of the closed-loop control model model into its internal representation (IR).  We
instrument this IR for each synthesis or verification scenario accordingly,
and use CBMC as an oracle to answer our queries.  CBMC itself relies on an
underlying SAT/SMT solver to address verification checks.  We~model the 
effects of finite precision arithmetic 
explicitly using CBMC's nondeterminism API (e.g., \texttt{nondet},
\texttt{CPROVER\_assume} intrinsic functions).  Our module is included into
CBMC~5.7 and is available for 
download.\footnote{\url{https://github.com/diffblue/cbmc/archive/cbmc-5.7.zip}}
We have further created a VirtualBox OVA image with our tool pre-installed,
including our MATLAB-generated benchmarks and benchmark shell script, 
with instructions to run all benchmarks and to reproduce our
experiments.\footnote{\note{\url{XXXXXXXXXX}}}
%
%\textcolor{red}{This paragraph needs to be updated by Elizabeth and Dario.}
%
%The entire implementation consists of $XXX$ and $XXX$ lines of C code in the multi-stage and abstract
%accelerator back-ends, respectively, $XXXX$ lines of C code for utility functions such as interval, 
%fixed- and floating-point arithmetic, as well as $1000$ lines of C++ code in the control synthesis
%module in CEGIS CBMC.  We evaluate \tool~using $XXX$ SISO control system
%benchmarks.

@@

Our benchmark suite consists of 19 case studies extracted from the control literature~\cite{acrobot,cstr,CHEN1979389,KOKOTOVIC198023,gajic2008optimal,Franklin15, maglev, converters, CTMS, gajic2008optimal}. These case studies comprise control models, which we represent in state space form as in \eqref{eq:ode1}. 
The models are time discretized, with sampling times ranging from 1s to 0.0001s. 
CEGIS initially starts with the coarsest discretisation (i.e., 1s sampling time), and if it fails to find a controller, reduces the sampling time. 
\note{we need to at least qualitatively mention input selection, initial conditions, and also safe sets. how much can you say? }

@@

that is a computer storage device used to read and write data on magnetic tapes. \\ 
The \textit{USCG cutter tampa heading angle} plant describes the heading angle dynamics of a Coast 
Guard cutter.

@@

Using the state-space models described in Section~\ref{exp:benchmarks}, 
the evaluation study has the following experimental goal:  
\begin{enumerate}
\item[] 
%[EG1] \textbf{(Performance)} 
Show that the multi-staged 
%and the abstraction-based 
CEGIS approach is able to generate finite precision digital controllers using fixed-point and floating-point arithmetic in a reasonable amount of time.
%\item[EG2] \textbf{(Sanity check)} Confirm the stability \note{also?} and safety of the synthesized controllers outside 
%of our model representation using MATLAB. 
%\note{Where is this done?}
\end{enumerate}

@@

We provide the results in Table~\ref{tab:results}, where: 
\textit{Benchmark} is the name of the corresponding benchmark; 
\textit{Order} is the number of continuous variables of the model; 
$\mathcal{F}_{\langle I_p,F_p \rangle}$, where $I_p$ and $F_p$ indicate the integer and fractional parts,  
and $\mathcal{F}_{\langle E_p,M_p \rangle}$, where $E_p$ and $M_p$ indicate the exponent and mantissa part, 
are the fixed- and floating-point precisions used to model the plant, respectively; 
and \textit{Time} is the total time required to synthesize a controller for the given plant.    
Timeouts \note{above: failures} are indicated by~\xmark.  
The precision for the controller, $\mathcal{F}_{\langle I_c,F_c \rangle}$, is
chosen to be $I_c = 8$, $F_c = 8$ for fixed-point, whereas  
$\mathcal{F}_{\langle E_c, M_c \rangle}$ is chosen to be $W_c = 6$ and $F_c = 10$
for half-precision floating-point format.

@@

The mean run-time for the successful benchmarks is 225\,s for fixed-point and 125\,s for floating-point implementations, respectively.  
The median run-time for the successful benchmarks is 15.7\,s for fixed-point and 17.8\,s for floating-point implementations, respectively.
We consider these times to be short enough to be of practical use to control engineers, and thus affirm the validity of 
%EG1. 
the objective we have raised.

\smallskip

@@

In this article we have presented a selection of case studies for which we have been able to automatically synthesise safe controllers. 
In the full experimental environment we have included further case studies that we have been unable to synthesise controllers for. 
This is in many cases due to timeouts, 
especially when the set completeness threshold is too large, 
or when a controller simply may not exist for a benchmark at a given time discretisation and required controller precision. 
Yet another source of incompleteness is the inability of the {\sc synthesize} phase to employ a large-enough precision for the plant model.

@@

We identify the following factors as potential limits to the generality of the obtained results.

\item[] 
\textit{Benchmark selection:} We report an assessment of both our approaches
over a diverse set of real-world benchmarks.  Nevertheless, this set of
benchmarks is limited within the scope of this paper and the performance may
not generalize to other benchmarks.
%\note{is this where we attempt asserting EG2?}

\item[]
\textit{Plant model precision and discretization heuristics:} Our algorithm to
select suitable finite-precision arithmetic to model the plant behavior 
increases the precision by 8 bits at each step,  
in order to be compliant with the CBMC type API. 
Similarly, for time-discretization, we try a set of pre-defined time discretisations.
This works sufficiently well for our benchmarks, but
performance may suffer in some cases, for example if the completeness
threshold is high.

@@

\end{itemize}

@@

We have presented an automated approach to synthesize provably-correct digital state-feedback controllers that ensure both stability and safety over state-space models.   
%The first approach relies on unfolding of the closed-loop model dynamics up to a completeness threshold, while the second
%one applies abstraction-refinement and acceleration to increase speed, whilst retaining soundness.
%
Our approach is novel within the digital control literature: 
we provide a fully automated synthesis method that is algorithmically and numerically sound,
considering various error sources in the implementation of the digital control algorithm and in the computational modeling of plant dynamics.
%
Our experimental results show we are able to synthesize fully automatically safe controllers for diverse benchmarks from the control literature, 
within a reasonable amount of time.  
%
%\note{Let us collect here the main extensions we regard to be relevant: \\ 
%models with partial observations; \\ 
%implementation with SMT; \\ 
%continuous-time; \\ 
%more complex specs; \\ 
%\ldots }
